# Appendix {.unnumbered .appendix}

## Software requirements {.unnumbered}

```{r}
#| label: r-version
#| echo: false
renv <- jsonlite::read_json("renv.lock")
```

This book was built using R version `r renv[[1]]$Version`. This book is built with [Quarto](https://quarto.org/), using version number `r quarto::quarto_version()`. All R packages required to build this book can be found in the following table. Note that this table contains all packages required to create the book, not just those required for the examples.

```{r}
#| label: tbl-package-versions
#| tbl-cap: "R packages and version numbers"
#| echo: false
pkgs <- unlist(lapply(renv[[2]], function(x) x[["Package"]]))
pkg_vs <- unlist(lapply(renv[[2]], function(x) x[["Version"]]))
nr <- ceiling(length(pkgs)/3)
pkg_df <- tibble::tibble(
  Package = pkgs,
  Version = pkg_vs
)
kableExtra::kbl(
  list(pkg_df[1:nr,], pkg_df[(nr+1):(2*nr),], pkg_df[(2*nr+1):length(pkgs),]),
  booktabs = TRUE,
  longtable = TRUE
) |> 
  kableExtra::kable_styling(font_size = 7)
```

