---
filters:
  - line-highlight
execute: 
  freeze: auto
---

```{r}
#| echo: false
#| eval: true
#| file: R/options.R
```

```{r}
#| echo: false
#| eval: true
#| file: R/load_font_awesome.R
```

```{r}
#| echo: false
#| eval: true
#| file: R/social_caption.R
```

# Nobel Prize Laureates: positioning text and parameterising plots {#sec-nobel}

## Data

* Originally used as a TidyTuesday data set back in 2019. Since that data is a little bit outdated. 
* Use API
* Different categories, start with Physics

[@nobel_data]

```{r}
#| label: nobel-load-data-show
#| eval: false
#| echo: true
nobel_physics <- read.csv("http://api.nobelprize.org/2.1/laureates?limit=250&nobelPrizeCategory=phy&format=csv&csvLang=en")
write.csv(nobel_physics, "data/nobel_physics.csv", row.names = FALSE)
```

```{r}
#| label: nobel-load-data-hide
#| output: false
#| echo: false
#| eval: true
nobel_physics <- readr::read_csv("data/nobel_physics.csv")
```

dataw wrangling

```{r}
physics_data <- nobel_physics2 |> 
  dplyr::select(name, gender, category, year)
```

## Exploratory work

### Data exploration

### Exploratory sketches

## Preparing a plot

### Data wrangling

### Packages and functions

### The first plot

## Advanced styling

### Colours

### Text and fonts

### Adjusting themes

## Reflection